Potential Factors of Decreasing Birth Rates in the US since the Great Recession and Beyond

### 2024-02-08 outputs/paper/paper.qmd
Update qmd file

# Plotting the data
ggplot(data, aes(x = year, y = brate_all)) +
  geom_line(color = "blue") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  annotate("text", x = 2007, y = 80, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
  labs(x = "Year", y = "Births per 1,000 women age 15-44",
       title = "Trend in US Birth Rates",
       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
  theme_classic()

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-trend-brate-graph
#| fig-cap: "Trends in Birth Rates by Population Subgroup"
# Load necessary libraries
library(ggplot2)
library(gridExtra)

@@

# Plotting the first line graph (Graph A)
graph_A <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = brate_1519, color = "Age 15-19"), size = 1.2) +
  geom_line(aes(y = brate_2024, color = "Age 20-24"), size = 1.2) +
  geom_line(aes(y = brate_2529, color = "Age 25-29"), size = 1.2) +
  geom_line(aes(y = brate_3034, color = "Age 30-34"), size = 1.2) +
  geom_line(aes(y = brate_3539, color = "Age 35-39"), size = 1.2) +
  geom_line(aes(y = brate_4044, color = "Age 40-44"), size = 1.2) +
  geom_text(aes(label = "Age 15-19", x = 2020, y = brate_1519), hjust = 0, vjust = -0.5, color = "blue", family = "Times") +
  geom_text(aes(label = "Age 20-24", x = 2020, y = brate_2024), hjust = 0, vjust = -0.5, color = "orange", family = "Times") +
  geom_text(aes(label = "Age 25-29", x = 2020, y = brate_2529), hjust = 0, vjust = -0.5, color = "grey", family = "Times") +
  geom_text(aes(label = "Age 30-34", x = 2020, y = brate_3034), hjust = 0, vjust = -0.5, color = "yellow", family = "Times") +
  geom_text(aes(label = "Age 35-39", x = 2020, y = brate_3539), hjust = 0, vjust = -0.5, color = "cyan", family = "Times") +
  geom_text(aes(label = "Age 40-44", x = 2020, y = brate_4044), hjust = 0, vjust = -0.5, color = "green", family = "Times") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5), labels = function(x) paste(x, "\n")) +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
  labs(y = "Births per 1,000 women in relevant population subgroup",
       title = "A: Five-year age group",
       caption = "Note: Birth rates by age group, race and ethnicity, and marital status are gathered from CDC Vital Statistics Births Reports. Birt rates by Hispanic subpopulation, mother's education are calculated by the original author's NCHS Vital Statistics birth microdata, SEER population data, and the Current Population Survey.") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"))

```{r}
# Plotting the second line graph (Graph B)
graph_B <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = brate_hisp, color = "Hispanic"), size = 1.2) +
  geom_line(aes(y = brate_blacknh, color = "Black, non-Hispanic"), size = 1.2) +
  geom_line(aes(y = brate_whitenh, color = "White, non-Hispanic"), size = 1.2) +
  geom_text(aes(label = "Hispanic", x = 2020, y = brate_hisp), hjust = 0, vjust = -0.5, color = "grey", family = "Times") +
  geom_text(aes(label = "Black, non-Hispanic", x = 2020, y = brate_blacknh), hjust = 0, vjust = -0.5, color = "orange", family = "Times") +
  geom_text(aes(label = "White, non-Hispanic", x = 2020, y = brate_whitenh), hjust = 0, vjust = -0.5, color = "blue", family = "Times") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5), labels = function(x) paste(x, "\n")) +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
  labs(title = "B: Race and ethnicity (ages 15-44)") +
  theme_classic()
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"))

>>>>

# Plotting the data
ggplot(data, aes(x = year, y = brate_all)) +
  geom_line(color = "blue") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  annotate("text", x = 2007, y = 70, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
  labs(x = "Year", y = "Births per 1,000 women age 15-44",
       title = "Trend in US Birth Rates",
       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.", family = "Times") +
  theme_classic()

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-bratetrend-graph
#| fig-cap: "Trends in Birth Rates by Population Subgroup"

#### Figure 2 ####

#### ####

# Load necessary libraries
library(ggplot2)
library(gridExtra)

@@

# Plotting the first line graph (Graph A)
graph_A <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = brate_1519, color = "blue"), size = 1.0) +
  geom_line(aes(y = brate_2024, color = "orange"), size = 1.0) +
  geom_line(aes(y = brate_2529, color = "grey"), size = 1.0) +
  geom_line(aes(y = brate_3034, color = "yellow"), size = 1.0) +
  geom_line(aes(y = brate_3539, color = "cyan"), size = 1.0) +
  geom_line(aes(y = brate_4044, color = "green"), size = 1.0) +
  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_1519, label = "Age 15-19"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_2024, label = "Age 20-24"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_2529, label = "Age 25-29"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_3034, label = "Age 30-34"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_3539, label = "Age 35-39"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_4044, label = "Age 40-44"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  annotate("text", x = 2007, y = 120, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = NULL) +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
  labs(x = NULL, 
       y = 'Births per 1,000 women in relevant population subgroup',
       title = "A: Five-year age group",
       caption = NULL,
       family = "Times") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times"),
        plot.title = element_text(size = 10, family = "Times"),
        legend.position = "none")

# Plotting the second line graph (Graph B)
graph_B <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = brate_hisp, color = "grey"), size = 1.2) +
  geom_line(aes(y = brate_blacknh, color = "orange"), size = 1.2) +
  geom_line(aes(y = brate_whitenh, color = "blue"), size = 1.2) +
  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_hisp, label = "Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_blacknh, label = "Black, non-Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
  geom_text(data = data[data$year == 2020,], aes(x = year, y = brate_whitenh, label = "White, non-Hispanic"), hjust = -0.1, vjust = -0.5, size = 2, family = "Times") +
  geom_vline(xintercept = 2007, linetype = "dotted") + annotate("text", x = 2007, y = 120, label = "2007", vjust = -0.5, hjust = 0.5, color = "black", family = "Times") +
  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") + scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20), labels = NULL) + scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) + labs(x = NULL, y = NULL,
       title = "B: Race and ethnicity (ages 15-44)",
       caption = "Birth rates by age group, race, and ethnicity, and marital status gathered from CDC Vital Statistics Births Reports. Birth rates by Hispanic subpopulation, mother's education are calculated by the original author's using NCHS Vital Statistics birth micro data, SEER population data, and the Current Population Survey.") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = -45, hjust = 1, family = "Times"),
        plot.title = element_text(size = 10, family = "Times"),
        legend.position = "none")
